Exploring probability distributions for bivariate temporal granularities

### 2019-12-08 main.Rmd
cyclic calendar algebra update

# Computing cyclic time granularities {#sec:order}

The hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities. We shall use the notion of a hierarchy table and order to define them.

**Definition:  Order** of a linear granularity can be comprehended as the level of graininess associated with a linear granularity. If we consider two linear granularities G and H, such that G is finer than or groups into H (by Definitions \ref{def:finerthan}, \ref{def:groupsinto}), then H is of higher order than G.

**Notation:  Hierarchy table** Let $H_n: (G, C, k)$ be a hierarchy model containing n linear granularities. $G_{l}$ represents the linear granularity of order l and $k(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the circular granularity that relates linear granularity of order l and m, $\forall l,m \in n, l<m$.

In a hierarchy table, linear granularities are arranged from lowest to highest order of linear granularities.

@@

We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities and those concerning a single level as **single-order-up** granularities. Let us look at few calendars to see examples of single and multiple order-up granularities.

@@

|  G	        |C                 |k
|------------	|------------------|------------|
|  kin      	|kin-of-uinal      |  20        |
|  uinal    	|uinal-of-tun      |  18  	    |
|  tun      	|tun-of-katun      |  20	      |
|  katun    	|katun-of-baktun   |  20        |
|  baktun   	| 1                |	1         |

>>>>

# Cyclic calendar algebra {#sec:cyclic-calendar}

In \autoref{sec:}, we discussed how we can define cyclic granularities by extending the concept of linear granularities defined in @Bettini1998-ed. In this section, we will see how to obtain cyclic granularities through an algebraic representation of other cyclic granularities. This is similar to calendar algebra in @Ning2002-tf, where linear granularities are generated from the bottom linear granularity or from those that are already generated. Since, our method caters to computation of cyclic granularities, we shall refer to it as "cyclic calendar algebra".

The cyclic calendar algebra consists of two kinds of operations: one that requires computing multiple order-up from single order-up cyclic granularities and the other that require the reverse. The hierarchical structure of time creates a natural nested ordering which can produce **single-order-up** or **multiple-order-up** granularities. We refer to granularities which are nested within multiple levels as **multiple-order-up** granularities and those concerning a single level as **single-order-up** granularities. We shall use the notion of a hierarchy table and order to define them.

**Notation:  Hierarchy table** Let $H_n: (G, C, P)$ be a hierarchy model containing n linear granularities. $G_{l}$ represents the linear granularity of order l and $P(l,m)$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the cyclic granularity that relates linear granularity of order l and m, $\forall l,m \in n, l<m$.  **Order** of a linear granularity can be comprehended as the level of graininess associated with a linear granularity. For example, if we consider two linear granularities $G$ and $H$, such that $G$ is finer than or groups into $H$, then $H$ is of higher order than $G$. In any hierarchy table, linear granularities are arranged from lowest to highest order of linear granularities. Let us look at some examples of single and multiple order-up granularities.

@@

@@

|  G	        |C                 |P
|------------	|------------------|------------|
|  kin      	|kin-of-uinal      |  20        |
|  uinal    	|uinal-of-tun      |  18  	    |
|  tun      	|tun-of-katun      |  20	      |
|  katun    	|katun-of-baktun   |  20        |
|  baktun   	| 1                |	1         |
